### 2023-11-15 supplementary_BiSC_outdoor_AP.tex

\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}
\usepackage{caption}
% Set page size and margins
% Replace `letterpaper' with`a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{authblk} % this is for multiple authors 
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{cleveref}

% Configure for "Fig.":
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Fig.}{Figs.}

% reference configuration
\usepackage[authoryear,round]{natbib}
\bibliographystyle{apalike} % for an APA-like style

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Tittle of the paper --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\textbf{Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Authors and affiliations --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% --- Authors --- %%%
\author[1, 2, 3]{Alan Domínguez}
\author[1, 3, 4]{Payam Dadvand}
\author[1]{Marta Cirach}
\author[1]{Bruno Raimbault}
\author[1]{Toni Galmes}
\author[1]{Karl Samuelsson}
\author[1, 2, 3]{Mark Nieuwenhuijsen}
\author[1, 2, 3]{Jordi Sunyer}
\author[1, 2, 3]{Xavier Basagaña}
\author[1, 3]{Ioar Rivas}

%%% --- Affiliations --- %%%
\affil[1]{Barcelona Institute for Global Health (ISGlobal), Barcelona, Spain.}
\affil[2]{Universitat Popmpeu Fabra (UPF), Barcelona, Spain.}
\affil[3]{CIBER Epidemiología y Salud Pública (CIBERESP), Madrid, Spain.}
\affil[4]{London School of Hygiene and Tropical Medicine ( LSHTM), London, UK.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

% Index table with to put order to supplementary materials

\section{Predictor variables}
\section{}

\section{Model evaluation metrics}

The following metrics were used to evaluate the LUR model for all the air pollution assessed.

% --- R2 --- %
\begin{equation}
R^2 = \left( \frac{\sum (\hat{y}_i - \bar{y})^2}{\sum (y_i - \bar{y})^2} \right)
\end{equation}

% --- root mean squared error (RMSE) --- % 
\begin{equation}
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)^2}
\end{equation}

% --- mean bias (MB) --- % 
\begin{equation}
\text{Mean Bias} = \frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)
\end{equation}

% --- correlation coefficient (r) --- %
\begin{equation}
r = \frac{\sum (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum (x_i - \bar{x})^2 \sum (y_i - \bar{y})^2}}
\end{equation}
\vspace{0.5 cm}

For the DM models we evaluate their performance using a different set of metrics as following:

% --- Percentage of observed-modelled pairs (n) --- %
\begin{equation}
n = \frac{N_{\text{pairs used}}}{N_{\text{total pairs}}} \times 100
\end{equation}

% --- Factor of Two (FAC2) --- % 
\begin{equation}
\text{FAC2} = \frac{1}{n} \sum_{i=1}^{n} U,
\end{equation}
%Where \( U \) is 1 if \( 0.5 \leq \frac{M_i}{O_i} \leq 2 \) and 0 otherwise.

% --- Mean Bias (MB) --- % 
\begin{equation}
\text{MB} = \frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)
\end{equation}

% --- Mean Gross Error (MGE) --- % 
\begin{equation}
\text{MGE} = \frac{1}{n} \sum_{i=1}^{n} |O_i - M_i|
\end{equation}

% --- Normalized Mean Bias (NMB) --- % 
\begin{equation}
\text{NMB} = \frac{\sum_{i=1}^{n} (M_i - O_i)}{\sum_{i=1}^{n} O_i}
\end{equation}

% --- Root Mean Square Error (RMSE) --- % 
\begin{equation}
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)^2}
\end{equation}

% --- Correlation Coefficient (r) --- % 
\begin{equation}
r = \frac{\sum_{i=1}^{n} (O_i - \bar{O})(M_i - \bar{M})}{\sqrt{\sum_{i=1}^{n} (O_i - \bar{O})^2 \sum_{i=1}^{n} (M_i - \bar{M})^2}}
\end{equation}
\vspace{0.5cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- References --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{references}

\end{document}