### 2024-02-19 manuscript_BiSC_outdoor_AP.tex
finished paper for co-authors

\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}
\usepackage{caption}
% Set page size and margins
% Replace `letterpaper' with`a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{authblk} % this is for multiple authors 
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{multirow}

% reference configuration
\usepackage[authoryear,round]{natbib}
\bibliographystyle{apalike} % for an APA-like style
\usepackage{amsmath}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Title of the paper --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\textbf{\Large Development and comparison of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Authors and affiliations --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% --- Authors --- %%%
\author[1, 2, 3]{\small Alan Domínguez}
\author[1, 2, 3]{Payam Dadvand}
\author[1, 2, 3]{Marta Cirach}
\author[4]{Gustavo Arevalo}
\author[4]{Lluís Barril}
\author[1, 2, 3, 5]{María Foraster}
\author[1, 2, 3]{Mireia Gascon}
\author[1]{Bruno Raimbault}
\author[1]{Toni Galmes}
\author[1, 2, 3]{Laura Goméz-Herrara}
\author[1, 2, 3]{Cecilia Persavento}
\author[1, 6]{Karl Samuelsson}
\author[4]{Jose Lao}
\author[7]{Teresa Moreno}
\author[7]{Xavier Querol}
\author[1, 2, 3]{Cathryn Tonne}
\author[1, 2, 3]{Mark Nieuwenhuijsen}
\author[1, 2, 3]{Jordi Sunyer}
\author[1, 2, 3, *]{Xavier Basagaña}
\author[1, 2, 3, *]{Ioar Rivas}

%%% --- Affiliations --- %%%
\affil[1]{\small Barcelona Institute for Global Health (ISGlobal), Barcelona, Spain.}
\affil[2]{Universitat Pompeu Fabra (UPF), Barcelona, Spain.}
\affil[3]{CIBER Epidemiología y Salud Pública (CIBERESP), Madrid, Spain.}
\affil[4]{Barcelona Regional}
\affil[5]{Blanquerna School of Health Science, Universitat Ramon Llull (URL), Barcelona, Spain.}
\affil[6]{Department of Building Engineering, Energy Systems and Sustainability Science, Faculty of Engineering and Sustainability Science, University of Gävle, Gävle, Sweden}
\affil[7]{Institute of Environmental Assessment and Water Research (IDAEA-CSIC), Barcelona, Spain.}
\affil[*]{Join last authors}

% --- removing date from manuscript --- % 
\date{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Abstract --- %%%
%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Air pollution is the leading environmental risk factor for human health worldwide linked to a wide range of adverse health outcomes. A detailed and comprehensive assessment of outdoor air pollution exposure considering both spatial and temporal variability in urban areas is essential to identify critical windows of susceptibility
during developmental periods over the lifespan to traffic-related air pollution (TRAP). Within the framework of the Barcelona Life Study Cohort (BiSC), we developed Land use regression (LUR), dispersion (DM), and hybrid (HM) models to estimate outdoor air pollution exposure at residential addresses for several TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn). The overall aim is to obtain accurate weekly estimates of traffic-related air pollution, intended to be used in the exposure assessment for a cohort of pregnant women in the city of Barcelona. Two separate monitoring campaigns were conducted: the "BiSC-home" campaign took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} between 2018-2021. Additionally, a second campaign was conducted, named "BiSCAPE", and included four distinct measurement periods between 2021-2022. LUR and DM were constructed using conventional state-of-the-art techniques, while HM was designed by harnessing Random Forest algorithm, integrating outputs from DM with explanatory variables from the LUR model, monitoring stations and detailed traffic counts. Model performance was evaluated using a cross-validation procedure for LUR and HM, and by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations. The models demonstrated moderate to good performance for LUR, DM and HM. HM outperformed LUR and DM models for NO\textsubscript{2} (R\textsuperscript{2} = 0.64), PM\textsubscript{2.5} (R\textsuperscript{2} = 0.66), and BC (R\textsuperscript{2} = 0.86). LUR models show better performance for PM\textsubscript{2.5 - Fe} (R\textsuperscript{2} = 0.89), PM\textsubscript{2.5 - Cu} (R\textsuperscript{2} = 0.87), PM\textsubscript{2.5 - Zn} (R\textsubscript{2} = 0.85). Correlations between the air pollutant estimates for the different models were moderate to strong for both long-and-short term exposures. Long-term exposure (entire pregnancy) at the residential addresses exhibited correlations ranging from rs = 0.48 to 0.92, while for short-term exposure (weekly exposure) varied from rs = 0.37 to 0.90. Our work emphasizes the importance of employing diverse exposure assessment methods to evaluate the exposure to traffic-related air pollutants and to disentangle differences in effect estimates during pregnancy. The exposure estimates from these different models can be used in the BiSC cohort to potentially identify critical windows of susceptibility to air pollution during pregnancy. \\

\textbf{Keywords:} Air Pollution; Traffic-related air pollutants; Exposure assessment; Spatiotemporal models.

\end{abstract}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Introduction --- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Air pollution stands as the primary environmental risk factor for human health and the fourth most deadly health risk worldwide, playing a significant role in the global burden of disease \cite{cohen2017, he2020state}. Exposure to ambient air pollution has been linked with many adverse health outcomes \cite{boogaard2022, guxens2022hei, haddad2023}. In this context, urbanization has been playing a major role in increasing air pollution exposure as cities grow and more population moves to urban areas \cite{nieuwenhuijsen2016}. These trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas during critical developmental stages \cite{boogaard2022}.

Over the past few decades, several modeling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution \cite{hoek2017, di2019no2, di2019pm25, stafoggia2019, stafoggia2020}. Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches \cite{gulliver2015, dehoogh2014}. LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that work as proxies for traffic-related air pollution \cite{briggs1997, hoek2008}. In contrast, DMs rely on mathematical simulations to estimate air pollution levels. These predictions are based on data from emissions sources, chemical transformation models, pollutant dispersion models, topography, roughness, and meteorological data \cite{hoek2017}. Although these methods have been extensively used they come with inherent limitations related to lack of temporal and spatial variation. To overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}.

While significant advancements have been made in spatiotemporal modeling of air pollution over large-scale domains (multi-country and national domain models), these models often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}. This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to capture the differences in exposure for the population living in relatively small areas (i.e. city compared to country-level) to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potentially lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.

Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that make difficult an accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}. In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees-based models \cite{liu2022}. This family of models is often used due to its ability to model potential nonlinear interaction between predictor variables \cite{liu2022treebased}. Among the different ML algorithms, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}. Furthermore, the COVID-19 pandemic has introduced unprecedented challenges for air pollution modelling and exposure assessment. These challenges encompass changes in air pollution patterns, constraints on data collection due to health and safety restrictions (e.g. lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}.

This study was developed within the framework of the Barcelona Life Study Cohort (\href{https://projectebisc.org/en/home/}{BiSC}), a pregnancy cohort that enrolled 1,080 pregnant women in their first trimester to evaluate the impact of exposure to air pollution during pregnancy, fetal and early life development of their offspring. We developed LUR, DM, and HM models to estimate outdoor air pollution exposure at residential addresses of the BiSC participants during their pregnancies. LUR and DM were constructed using conventional techniques, while HM was designed by harnessing RF algorithm, integrating outputs from DM with explanatory variables from the LUR model. The overall aim is to obtain accurate weekly estimates for NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn), intended to be used in the prenatal exposure assessment for the BiSC cohort.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Material and methods--- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Study domain and Monitoring Campaigns --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study domain and monitoring campaigns}

Barcelona (\textbf{\cref{fig1}}), is one of the most populated and dense cities in Spain, with 1,656,725 inhabitants ($\approx16,255.4$ inhabitants/km\textsuperscript{2}). The Barcelona Metropolitan Area (AMB) is located on the Mediterranean coast of the Iberian peninsula, protected by the Collserola mountain range and delimited by two delta rivers (Besòs and LLobregat). The city has a Mediterranean climate, the wind patterns are dominated by a breeze that blows from the sea during the daytime and from the land during night-time. Barcelona is one of the cities in Europe with the highest vehicle traffic density (6,000 vehicles/km\textsuperscript{2}) \cite{casallas2018}. Air pollution measurements were conducted during two separate monitoring campaigns. The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021. The second campaign, named "BiSCAPE", was carried out following the European Study of Cohorts for Air Pollution Effects (ESCAPE) project guidelines \cite{ESCAPE2010}, and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022. During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, Cornellà de Llobregat, Hospitalet de Llobregat and Sant Just Desvern) (\textbf{\cref{fig1}}). These monitoring campaigns were used in the development of the LUR and HM. Air pollution levels were measured using specialized  equipment for each pollutant. NO\textsubscript{2} was measured using Gradko NO\textsubscript{2} diffusion tubes. PM\textsubscript{2.5} and its chemical organic components were measured using a 37mm Polytetrafluroethylene - Teflon® - filters collected using a BGI-400 pump (working at 4 L/min) and a PCIS impactor, and BC measurements were conducted using a MicroAeth®.

% Put the figure text closer to the Figure 1
\captionsetup[figure]{skip=0pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/figure_campaigns.png}
\caption{Study domain and spatial distribution of the monitoring campaigns}
\label{fig1}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- LUR model --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LUR models}

LUR models for NO\textsubscript{2} were developed using only data from the locations that had measured home-outdoor (BiSC-home campaign) levels for both sampling campaigns (i.e., we did not use data from participants who had data available for only one campaign, either first or third trimester). Almost half of the sites were measured before the COVID-19 pandemic started and half of the sites were during and after the pandemic period and its resulting restrictions. For PM\textsubscript{2.5}, BC  and PM\textsubscript{2.5} constituents, models were developed using BiSCAPE data from the 34 sites. To develop the LUR models, spatial variables that better represent the air pollutant concentration distribution across the study area were created. Following the ESCAPE guidelines \cite{ESCAPE2010, eeftens2012, beelen2013},  predictor variables were selected from three broad categories: traffic, land use, and urban configuration. Various buffer sizes were created, i.e, 25m, 50m, 100m, 300m, 500m, and 1000m to capture the area-specific predictor variable at every location. Predictor variables were calculated using specific coordinates for each site. We applied a supervised stepwise process in which we sequentially added a predictor variable to a linear regression model using pollutant concentrations as the independent variable to obtain the best possible model \cite{eeftens2012, beelen2013}. This procedure was followed to build all the specific pollutant models. Previously to model development, we specified the direction expected of the effect that each predictor has in the model. For all predictor variables, univariate regression models with annual average air pollution concentration (NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn)) were conducted, models with the highest adjusted explained variance (adjusted R\textsuperscript{2}) and the right direction of the effect were selected. We then repeated the process using the previous model and all the remaining predictor variables. Details of the LUR modeling and predictor variables can be found in the \textbf{supplementary materials, Table 1}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Dispersion model --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dispersion models}

An emission inventory was compiled for the period 2018-2021, which included data from diverse emission sources such as vehicular traffic, household activities, industrial plants, airports, harbors, factories, and the commercial sector. The Atmospheric Dispersion Modelling System software (ADMS-Urban) from Cambridge Environmental Research Consultants (CERC) was employed to calculate the concentration of pollutants like NO\textsubscript{2}, PM\textsubscript{2.5}, and BC originating from these sources. ADMS-Urban is a specialized Gaussian model designed for analyzing pollutant emissions in urban and metropolitan areas, particularly at road level \cite{mchugh1997adms}. ADMS-Urban employs various algorithms to assess the chemical transportation and dispersion of pollutants, factoring in ground-level turbulence and turbulence induced by surrounding buildings \cite{stocker2012adms}. The model also features a meteorological pre-processing system that calculates fluxes and turbulence, enabling it to model airflow over complex terrains.For enhanced accuracy, the dispersion model integrates data from a three-dimensional building and street width model from the Institut Cartogràfic i Geològic de Catalunya \cite{stoter2016state, 3Dcity}. This data aids in the calculation of the street canyon effect, a phenomenon characterized by tall buildings situated adjacent to heavily trafficked roads, which hinders atmospheric dispersion, leading to higher and relatively consistent pollutant concentrations compared to areas unaffected by the canyon effect. The model also utilized meteorological data including wind speed, direction, air temperature, cloud cover, solar radiation, relative humidity, precipitation rate, and Planetary Boundary Layer Height (PBLH). This meteorological data was collected from the Zona Universitària station in Barcelona, managed by the Servei Meteorològic de Catalunya \cite{xema2013}. To incorporate background pollution levels at a regional scale, data from Copernicus Atmosphere Monitoring Service Information (CAMS) was used. Background air pollution data has been extracted from the CAMS reanalysis ensemble model for 2018, and from CAMS ensemble forecast model for 2019-2020 \cite{cams2020, franceinstitut}. ADM results were evaluated considering the locations of XVPCA stations, to evaluate the model performance on those locations. The model was calibrated to fit the model results to the data measured in each XVPCA station. In addition, bias-adjustment techniques have been used to fit the model data to the observations through hourly bias spatial interpolation and generation of hourly maps. After that, data were processed  by cluster analysis techniques to obtain a mean cluster map. Each bias map by pollutant was associated to a cluster map class and an hourly-varying profile. Further details regarding the ADMS-Urban model and the variables used can be found in \textbf{supplementary materials}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Hybrid model --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hybrid models}
For the hybrid models, we sought to improve their predictive power by integrating comprehensive data from multiple information sources and leveraging ML. NO\textsubscript{2}, PM\textsubscript{2.5}, PM\textsubscript{2.5} constituents, and BC were largely based on daily concentrations from dispersion models, summarized as weekly averages, and spatial predictor variables from the LUR models. For a more comprehensive approach, we integrated additional data from the Catalan Air Pollution Monitoring and Forecasting Network (XVPCA) for Barcelona and surrounding areas \cite{xarxa2012}. Daily air pollution concentration levels were interpolated using the Inverse Distance Weighting (IDW) method, which assigns higher importance to nearest stations \cite{hoek2017methods}. The data were collected from 16 monitoring stations, including Ciutadella, Eixample, Gracia, Hospitalet, Ies Goya, Jardins, Observatori Fabra, Palau Reial, Plaza Universitària, Poblenou, Sagnier, Sants, St. Adria, St. Coloma, Vall Hebron, Zona Universitária \cite{xarxa2012}. While our dispersion model initially utilized meteorological and traffic data, the hybrid model broadened its scope by integrating daily traffic counts from 180 locations throughout the city \cite{trafficbcn}. These broader data helps us better understand the spatial and time-specific variations in urban air pollution \cite{pinto2020}. To further refine our insights into the influence of weather patterns on pollution, we sourced additional meteorological variables, such as wind speed and air temperature, from the Raval monitoring station managed by Servei Meteorològic de Catalunya \cite{xema2013}. We employed the RF algorithm to capture non-linear relationships and potential interactions between predictor variables and the response variable, utilizing \textit{the ranger package} in R implemented in the \textit{caret} framework \cite{wright2019}, a set of functions that attempt to streamline the process for creating predictive models \cite{caret2008}. Briefly, RF is a machine learning (ML) algorithm that constructs an ensemble of multiple decision trees \cite{breiman2001}. In this approach, each tree is formulated using a bootstrap sample of the data. During this process, every node within a tree represents the optimal split from a randomly selected subset of predictors. Predictions are then derived by averaging the outputs from all trees. In summary, by combining data from multiple sources, our hybrid models aim to provide more accurate and reliable predictions. The distribution of the air pollution monitoring stations, meteorological stations, and traffic count measures are shown in the (\textbf{\cref{fig1}}, \textbf{Panel D}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Performance assessment --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Performance assessment}
We evaluated model performance using various statistical approaches, validation datasets, and on different temporal resolutions. The LUR models were evaluated on an annual basis, DM was assessed both hourly and daily, and HM was evaluated on a weekly basis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- LUR model evaluation --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{LUR model evaluation}
Performance of the LUR models was assessed using \textit{k-fold cross-validation}. Specifically, we employed a 10-fold cross-validation technique. In this approach, the data was split into ten distinct sub-samples. For each iteration, the model was trained on \textit{k-1} folds and then validated using the leftover fold \cite{ziegel2003}. For every fold, we computed metrics such as the coefficient of determination \(R^2\) and the root-mean-square-error (RMSE). After each iteration, both the error and adjusted \(R\textsuperscript{2}\) were recorded. This process was repeated \(k-times\), and the average of these measures was reported as the indicator of the model validity. Details from the metrics used for model validation can be found in the \textbf{supplementary material}.

We carried out both internal and external validation for the NO\textsubscript{2} LUR model. Internal validation was conducted by applying a K-fold cross-validation analysis. The 10-fold cross-validation procedure and its significance have been detailed above. For the external validation of the NO\textsubscript{2} LUR model, we applied data from BiSC-home participants' homes for those participants who had data available only for one measurement campaign, and therefore their data were not used for developing NO\textsubscript{2} LUR model. For PM\textsubscript{2.5}, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn}, and BC, internal validation was conducted using the \textit{leave-one-out-cross-validation (LOOCV)} process. For these models, we were not able to conduct external validation due to the lack of data availability.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Dispersion model evaluation --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{Dispersion model evaluation}
Briefly, once the input data has been introduced to the ADMS-Urban model, the calculation of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC concentration levels has been carried out. Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality monitoring stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others. Additionally, we compared the results of the model with data from the 4 campaigns carried out at BiSCAPE sites. In total, 98 periods were used to evaluate the performance of NO\textsubscript{2} model, 66 periods for PM\textsubscript{2.5}, and 93 periods for BC. The closer these values are, the better the model's performance. Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Hybrid model evaluation --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hybrid model evaluation}
We assessed the performance of each hybrid model by contrasting predictions with observations. Two distinct validation approaches were employed. Firstly, \textit{10-fold cross-validation (10-fold CV} was used to provide an internal validation measure, where the model was trained on 80\% of the data. Secondly, external validation was performed by allocating 20\% of the data as a test set to further evaluate the model's performance against previously unseen data. While both methods differ in their approach to partitioning the data, the fundamental distinction lies in how the data is utilized during the training and validation phases. The \textit{10-fold CV} method systematically rotates through the data, ensuring that each subset serves as a validation set once, while the external validation approach provides a straightforward assessment against a separate data portion. Additionally, we employed a grid search for hyperparameter tuning to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Comparison between model estimates --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison between model estimates}
Exposure estimates obtained from LUR, DM, and HM were compared at the BiSC participant's address level for long- and short-term exposure. We compare the distribution of estimated concentrations between models and pollutants and calculate Spearman correlation coefficients. Bland-Altmant plots were used to investigate the level of agreement between the different types of models. \\

All statistical analyses were performed in R version 4.3.1 \cite{Rstudio}. GIS predictor variables were computed using X software.

%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Results --- %%%
%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --- Distribution of the monitoring campaigns --- % 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Distribution of air pollution measurements}

Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants are shown in \Cref{table1}. NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 and 97.3 \(\mu \text{g/m}^3\) with a median of  37.5 \(\pm\) 15.8 \( \mu \text{g/m}^3\). PM\textsubscript{2.5} observed levels ranged between 5.01 and 35.8 \( \mu \text{g/m}^3\) with a median of 11.3 \(\pm\) 7.14 \(\mu \text{g/m}^3\), while BC levels showed less variation during the BiSCAPE campaign, ranging from 0.40 to 2.95 \(  \mu \text{g/m}^3\) with a median of 1.22 \(\pm\) 0.64 \( \mu \text{g/m}^3\). For PM\textsubscript{2.5} constituents  observed values during the BiSCAPE campaign varied substantially between elements with medians of 0.21 \(\pm\) 0.17 \(  \mu \text{g/m}^3\)  for  Fe,  6.36 \(\pm\) 4.61 \( \text{ng/m}^3 \) for Cu, and 28.2 \(\pm\) 27.8 \( \text{ng/m}^3 \) for Zn.

%%% ---------------------------------------------------------------- %%%
%%% Table 1. Outdoor air pollution concentration BiSC-home, BiSCAPE  %%%
%%% ---------------------------------------------------------------  %%%
\begin{table}[h!]
\centering
\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns (2018-2022)}
\label{table1} 
\begin{tabular}{llllll} % each one indicates a new column in the table 
\toprule
Campaign & Pollutant & Year  & median \(\pm\) IQR & min & max \\ 
\midrule
BiSC-home + BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 37.5 \(\pm\) 15.8 & 6.85 & 97.3 \\
\hline
\multirow{5}{*}{BiSCAPE} 
& PM\textsubscript{2.5} (\(\mu\)g/m\textsuperscript{3}) & 
2021-2022  & 11.3 \(\pm\) 7.14 & 5.01 & 35.8 \\ 
& BC (\(\mu\)g/m\textsuperscript{3}) & 
2021-2022 & 1.22 \(\pm\) 0.64 & 0.40 & 2.95  \\ 
& PM\textsubscript{2.5 - Fe} (\(\mu\)g/m\textsuperscript{3}) & 
2021-2022 & 0.21 \(\pm\) 0.17 & 0.07 & 0.77 \\ 
& PM\textsubscript{2.5 - Cu} (ng/m\textsuperscript{3}) & 
2021-2022 & 6.36 \(\pm\) 4.61 & 1.78 & 22.7 \\ 
& PM\textsubscript{2.5 - Zn} (ng/m\textsuperscript{3}) & 
2021-2022 & 28.2 \(\pm\) 27.8 & 6.71 & 273 \\ 
\bottomrule
\end{tabular}
\end{table}

The spatial distribution of the outdoor air pollution campaigns is illustrated in \Cref{fig1}. \textbf{Figure 1A} shows a high-density coverage in the central area of Barcelona, where the BiSC-home campaign was conducted. \textbf{Figure 1B} exhibit a similar pattern in the spatial distribution of the monitoring campaign, albeit with a reduced density of sampling locations compared to the BiSC-home campaign.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%             
% --- Model performance: LUR, DM, HM --- % 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Model performance: LUR, DM, HM}

\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using LUR, HM, and HM. The performance of the models varied substantially across different pollutants as shown in \Cref{fig2} and \Cref{fig3}. In general, LUR and HM models exhibit relatively good performance, while DM shows a more conservative performance evaluated by the $R^{2}$ and $RMSE$. The HM demonstrated moderate to good performance for NO\textsubscript{2} ($R^{2}_{\text{ 10-CV}} = 0.64$), PM\textsubscript{2.5} ($R^{2}_{\text{ 10-CV}} = 0.66$), and BC ($R^{2}_{\text{ 10-CV}} = 0.86$), with corresponding $RMSE_{\text{ 10-CV}}$ values of $7.48 \, \mu\text{g/m}^{3}$, $3.45 \, \mu\text{g/m}^{3}$, and $0.23 \, \mu\text{g/m}^{3}$, respectively. However, for PM\textsubscript{2.5} components, the performance was more conservative, with PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ 10-CV}} = 0.54$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ 10-CV}} = 0.70$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ 10-CV}} = 0.44$).In contrast, LUR models showed better performance in predicting PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ LOOCV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ LOOCV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ LOOCV}} = 0.85$) concentration levels in comparison to  NO\textsubscript{2} ($R^{2}_{\text{ LOOCV}} = 0.62$), PM\textsubscript{2.5} ($R^{2}_{\text{ LOOCV}} = 0.45$), and BC ($R^{2}_{\text{ LOOCV}} = 0.83$) exposure concentrations. Dispersion models, when compared with observations from the XVPCA monitoring stations were evaluated against observations from the XVPCA monitoring stations, yielded an $R^{2} = 0.39$ for NO\textsubscript{2}, $R^{2} = 0.26$ for PM\textsubscript{2.5}, and $R^{2} = 0.65$ for BC. A comparison between predicted and observed values for all the monitoring stations used in the DM is provided in the \textbf{supplementary materials}.

%%% --- Figure 2. Average exposure during the entire pregnancy period --- %%%
\captionsetup[figure]{skip=4pt} % Reduced space after caption
\begin{figure}[!h]
\includegraphics[width=0.75\textwidth]{figures/combined_figure1_v2.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
\label{fig2}
\end{figure}

%%% --- Figure 3. Average exposure during the entire pregnancy period --- %%%
\begin{figure}[!h]
\includegraphics[width=0.75\textwidth]{figures/combined_figure2_v3.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and HM (\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
\label{fig3}
\end{figure}

The maps reveals distinct patterns of variability in estimated outdoor air pollution exposure concentrations. LUR models shown higher variability in exposure concentrations during pregnancy, whereas DM exhibit less variability. HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.

%%% --- Table 2. Performance from validation for LUR, DM, HM --- %%%
\begin{table}[h!]
\centering
    \caption{Performance from validation for LUR, DM, and HM models.}
    \label{table2}
    \begin{tabular}{llcccccc}
    \toprule
         Model & Validation & \multicolumn{6}{c}{R\textsuperscript{2 \textsubscript{ Validation}}} \\
         \cmidrule(lr){3-8}
         & & NO\textsubscript{2} & PM\textsubscript{2.5} & BC & PM\textsubscript{2.5 - Fe} & PM\textsubscript{2.5 - Cu} & PM\textsubscript{2.5 - Zn} \\
    \midrule
         LUR & External validation BiSC-Home & 0.56 & - & - & - & - & - \\
         DM & External validation BiSCAPE & 0.65 & 0.22 & 0.56 & - & - & - \\
         HM & External validation split (test data) & 0.64 & 0.65 & 0.69 & 0.65 & 0.56 & 0.36  \\
    \bottomrule
    \end{tabular}
\end{table}

Results from validation of the LUR, DM, and HM models, conducted where information for validated the model was available are shown in \Cref{table2}. The validation of each model has been conducted using different validation datasets. NO\textsubscript{2} LUR model was validated using BiSC - home campaign data. NO\textsubscript{2}, PM\textsubscript{2.5} and BC from DM model were validated using data from BiSCAPE monitoring campaign. While HM models were validated using 20\% of data available initially before model training. Due to lack of data availability we were able to evaluate the performance of NO\textsubscript{2} for LUR models. The performance of the LUR model decreased in the external validation by 16\%. When validated DM against BiSCAPE measurements model performance increase for NO\textsubscript{2}, but decrease for PM\textsubscript{2.5} and BC. HM showed consistent results in performance using both approaches to validate the models for NO\textsubscript{2} and PM\textsubscript{2.5} models showed a similar performance that the one obtained using the \textit{10-CV}. While for BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} performance varied between validation approaches by -19.7\%, +20.37\%, -20.0\% , and -18.8\% respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Comparison between models --- %%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison of LUR, DM and HM estimates at address level} % broad title then change this

% --- Exposure concentrations --- %
\textbf{\Cref{fig4}} presents the distribution of estimated NO\textsubscript{2}, PM\textsubscript{2.5}, BC,  PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} concentrations from the LUR, DM, and HM at the address level. The data shown are estimated concentrations of pollutants during the entire pregnancy period (weekly estimates) for all the study period (2019 - 2021). In our analysis, estimated median concentration for NO\textsubscript{2} was higher in the HM ($36.7 \, \mu\text{g/m}^{3}$) in comparison to LUR ($36.4 \, \mu\text{g/m}^{3}$) and DM ($27.5 \, \mu\text{g/m}^{3}$). However, for PM\textsubscript{2.5} median concentration was higher for LUR ($16.7 \, \mu\text{g/m}^{3}$) compared to HM ($12.2 \mu\text{g/m}^{3}$) and DM ($11.8 \, \mu\text{g/m}^{3}$). This trend was also observed in BC estimates, where LUR's median concentration was $1.34  \mu\text{g/m}^{3}$, higher than HM's ($1.21 , \mu\text{g/m}^{3}$) and significantly greater than DM's ($0.76 , \mu\text{g/m}^{3}$). For PM\textsubscript{2.5} constituents like Fe, Cu, and Zn, a different pattern was observed, with HM presenting higher median concentrations than those observed in LUR models. Overall, while HM generally indicated higher median concentrations estimates, LUR models demonstrated greater variability with larger ranges across all pollutants assessed. Dispersion models predicted lower exposure concentration values in all pollutants than LUR and HM. \\\\

%%% -------------------------------------------%%%%
%%% --- Figure 4. Boxplot models comparison --- %%%
%%% ------------------------------------------%%%%
\begin{figure}[!h]
\includegraphics[width=0.8\textwidth]{figures/boxplot_all_models_estimates.png}
\caption{Distribution of estimated exposure to outdoor air pollution for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
\label{fig4}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --- Correlation between measures and model estimates --- %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 \textbf{(figure 5.A)}. Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 \textbf{(figure 5.B)}. Across the different models assessed, the correlation between the air pollutant estimates was moderate to strongly correlated for both long-and short-term exposure periods. Estimated concentrations of NO\textsubscript{2} ranged from 0.85 to 0.92. In terms of PM\textsubscript{2.5}, the correlation for estimated concentrations from LUR, DM, and HM varied from r = 0.54 to 0.66, while correlations for BC were stronger ranging from r = 0.64 to 0.88. For PM\textsubscript{2.5} constituents we observed higher correlation coefficients for PM\textsubscript{2.5 - Fe} in LUR and HM for both long and short term exposure. Estimates for PM\textsubscript{2.5 - Cu} and PM\textsubscript{2.5 - Zn} showed lower correlations for both methods compared to Fe constituent. Overall, NO\textsubscript{2} and BC from LUR, DM, and HM were strongly correlated, while for PM\textsubscript{2.5} and constituents the correlation was weaker. Highest correlation was found for pollutants estimated using DM than LUR and HM based pollutants. Lowest correlation between outdoor air pollutants was found for HM estimates.

% Here add the correlation plot (long and short term)
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!h]
\includegraphics[width=1.0\textwidth]{figures/correlation_fig_v2.png}
\caption{Correlation between traffic-related air pollutants for LUR, DM, HM in different windows of exposure \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
\label{fig5}
\end{figure}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --- Comparison/agreement between air pollution estimates --- %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{\Cref{fig6}} shown Bland-Altmant plots to assess the level of agreement across the concentrations ranges between the models. The limits of agreement, represented by the red dashed lines  contains the majority of the points with some outliers present in the plots indicating good agreement in the estimated concentration, but with potential disagreement at higher estimated concentrations, showing an upward trend when comparing the model`s estimates. Overall, when comparing outdoor air pollution estimates at address level the models show good agreement at lower and medium ranges, with low bias that is observed as the blue ticked line is close to 0 for almost all comparison between models.

%%% --- Figure. Comparison/agreement between moodel predictions estimates --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!h]
\includegraphics[width=0.7\textwidth]{figures/final_bland_altmant.png}
\caption{Bland-Altmant plots for comparing outdoor air pollution estimates for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, and Zn). using LUR, DM, and HM}}
\label{fig6}
\end{figure}
\vspace*{-\baselineskip}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}

% --- General --- % 
We successfully developed outdoor air pollution models (NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents) to assess 
exposure within the city of Barcelona for the period 2018-2021. Utilizing three different approaches: land use regression, dispersion modeling, and hybrid models were tailored to provide exposure estimates across various time windows releveant to our study population in the BiSC cohort, including entire pregnancy duration, specific trimesters, and weeks of pregnancy for each participant. The performance of these models was moderate to good, as determined by different evaluation metrics. The LUR models accounted for approximately 45\% to 89\% of the variance in pollutant levels, indicating their strength in capturing spatial variability in air pollution concentrations. HM models, which integrate a broader range of data inputs and modeling techniques, explained 44\% to 86\% of the variance. In contrast, DM models, focusing on physical and chemical processes affecting pollutants, explained a lower range of variability, between 26\% and 65\%. These findings align with the outcomes of similar air pollution modeling studies conducted across Europe and specifically within Barcelona \cite{fung2024, mila2023, criado2023, rovira2022, chen2020, dehoogh2018, dehoogh2013}. For both LUR and HM approaches, the selection of predictor variables demonstrated considerable variability across different pollutants, reflecting the influence of distinct emission sources. Notably, BC achieved the highest model performance across all methods, followed by NO\textsubscript{2}, PM\textsubscript{2.5}, and their constituents. This pattern highlights the critical role of source-specific factors in influencing the accuracy and reliability of pollution exposure estimates.\\

% --- NO2 models --- %
The performance of our LUR NO\textsubscript{2} model was higher than previous European models \cite{rivera2012} as well as earlier models specific to Barcelona \cite{criado2023}, indicating an enhanced capacity to capture the intra-urban variability of NO\textsubscript{2} concentrations levels. Notably, our LUR model outperformed a recently developed microscale LUR model, which was part of a data fusion method designed to minimize prediction uncertainty by integrating a LUR base map with hourly dispersion estimates from the CALLIOPE-Urban model within Barcelona \cite{criado2023}. The predictor variables of our model were predominantly traffic-related, and were selected due to their strong association with NO\textsubscript{2} emissions \cite{casallas2018}. Conversely, our DM model exhibits a lower performance compared to the CALLIOPE-Urban system, a regional-to-street-scale modelling framework that offers high-spatial-and-high-temporal-resolution (10m x 10m, hourly) by integrating the mesoscale airquality forecasting system CALIOPE (WRF-HERMES-CMAQ-BSS-DREAM8b) with an urban roadwide dispersion model (R-LINE) \cite{benavides2019}. Regarding our HM, where we integrated detailed information on traffic counts across the city of Barcelona, air pollution and meteorological data from monitoring stations, and estimates from the DM model through a RF algorithm, we achieved better performance compared to both our LUR and DM models. Most relevant predictor variables were estimates from the DM, monitoring station, traffic counts, solar radiation traffic-related spatial variables. We attribute the enhanced variance explanation by the HM to two main factors. Firstly, the incorporation of outputs from the DM model allows to improve the spatiotemporal variability in our model \cite{michanowicz2016, korek2017}, in addition traffic information allows to capture other emission sources not previously captured \cite{pinto2020} in our other two approaches. Secondly, the integration of the RF algorithm enables the model to account for non-linear relationships and potential interactions between predictor variables that the LUR and DM models did not fully exploit \cite{araki2018, chen2019}. Despite these advancements, a recent hybrid model developed for the entire region of Catalonia, which utilized information on air pollution measurements from monitoring stations, land use variables, and satellite remote sensing products, outperformed the performance of our HM for NO\textsubscript{2} \cite{mila2023}. This improvement is likely attributable to the use of data from the TROPOspheric Monitoring Instrument (TROPOMI), providing deeper insights into the vertical atmospheric distribution of NO\textsubscript{2} \cite{petetin2023}.\\

% --- PM25 particles --- % add predictor variables 
PM\textsubscript{2.5} have a moderate performance in LUR models similar to  previous models that aim to predict within-city variability in spanish cities \cite{rivera2012, wang2014} but lower than other multi-cities LUR models developed in Europe \cite{eeftens2012, gulliver2018}. Predictor variables used in the LUR model were mainly traffic, road density, and residential variables. We believe that the performance obtained from our LUR model can only explain a part of the variation of the  PM\textsubscript{2.5} concentrations across Barcelona. One potential cause that could partially explain the moderate performance of our LUR model is that PM\textsubscript{2.5} in Barcelona came from nine to ten major contribution sources that have been previously identified in two chemical species and source contribution studies for PM\textsubscript{2.5} \cite{in2021, amato2024}, we believe that our model was not able to take into account the variability of this sources since LUR models mainly use traffic-related variables to predict PM\textsubscript{2.5} concentrations levels, and previously have been described that LUR models are less effective to predict sources other than traffic \cite{dehoogh2014}. The performance of our DM model was less effective than the one obtained from the LUR model. One potential explanation could be the limited coarser spatial resolution from DM compared to LUR. Our lUR model uses observed data and land use characteristics to estimate air pollution levels, potentially offering a finer spatial resolution that better captures local variations, especially within urban settings \cite{hoek2017}. Since there were no previously available published DM models PM\textsubscript{2.5} developed in the city of Barcelona or at a Catalunya scale we were not able to compare the performance. The enhanced performance of our PM\textsubscript{2.5} HM model compared to our previous approaches can largely be attributed to the incorporation of comprehensive information during the model-building process and  the employment of the RF algorithm significantly increased our model's accuracy in predicting air pollution levels. This approach enabled us to account for additional sources of variability in PM\textsubscript{2.5} concentration levels, identifying emissions sources not captured by the other models. Additionally, our HM model demonstrated improved performance metrics compared to a previously developed RF model \cite{mila2023}. Regardless of these advancements in modeling PM\textsubscript{2.5} concentrations by our HM, it was found that recent European multi-stage HM models, utilizing RF algorithm for predicting daily PM\textsubscript{2.5} levels at a country-level with a high spatiotemporal resolution, exhibit superior performance metrics. These models integrated satellite Aerosol Optical Depth (AOD) measurements along with PM and meteorological data monitored at various locations and spatial predictors at the grid cell level, enhancing predictive accuracy beyond our HM model \cite{dehoogh2018, stafoggia2019, stafoggia2020, schneider2020}. This difference in model performance can be attributed to two main factors. First, the comprehensive exposure assessment period employed by these models, which utilized routine several years of data from routine monitoring stations to more effectively capture temporal trends in air pollution exposure levels. Second, the inclusion of AOD measurements provided additional insights into the background and regional variability of air pollution levels with higher correlations in Western Europe compared to the rest of the Region \cite{chen2024}.\\

% --- Black Carbon (BC) --- %
BC models have the highest performance across all pollutants, this was systematically observed in all our exposure assessment methods explored. 
The LUR-developed model has greater performance than previous BC models developed in Europe \cite{hoek2015, van2020}. However, there are only a few studies that aim to predict BC concentrations levels across Europe, this limited our comparison with other models. Our HM outperformed a previous hybrid European-wide model developed using large-scale satellite data, chemical transport models, and small-scale and land use predictors to predict BC \cite{dehoogh2018}, and a hybrid LUR model that 
to estimate BC and metal components across the metropolitan area of Pittsburgh, USA \cite{tripathy2019}. Predictor variables used for the development of LUR and HM models were mainly traffic, residential, and population density-related. We believe that the high performance in both LUR and HM is that BC levels proportionally reflect traffic-related process  \\

% --- PM2.5 constituents particles --- % 
LUR has a very good performance in predicting PM\textsubscript{2.5} constituents pollutants, while HM has a moderate performance. In both LUR and HM models, traffic-related variables emerged as significant predictors for PM\textsubscript{2.5} constituents, this was consistent with results obtained from stepwise and RF analysis. This aligns with previous evidence indicating the influence of traffic-related variables that explain a large proportion of the variability in the measured concentrations \cite{dehoogh2013, chen2020}. Additionally, industry-related variables appears as relevant predictors in predicting PM\textsubscript{2.5 - Cu} and \textsubscript{2.5 - Zn} in both LUR and HM, possibly reflecting emissions from industry sector \cite{belis2013, chen2020}. Our PM\textsubscript{2.5} constituents models outperformed in performance to those published in previous studies \cite{}. Differences in performance between LUR and HM models in predicting PM\textsubscript{2.5} constituents could be explained due to stability for long-term concentrations in PM\textsubscript{2.5} composition.\\

% --- Predicted concentrations ranges --- %

% --- Correlations and agreement between models --- %
Comparisons of air pollution estimates at the address level indicated that the correlations between the different models' estimates were moderate to strong for both long-term and short-term exposure periods. For NO\textsubscript{2}, PM\textsubscript{2.5} and BC we observed a positive correlation for both long-and-short exposures. Long-term correlations were strong for NO\textsubscript{2}, PM\textsubscript{2.5} and BC in comparison to short-term correlations. This is mainly due to a smoothing effect over the pregnancy period that could lead to more stable correlations between pollutants and also a seasonality effect between trimesters \cite{wang2018, morgan2023, hao2023}. This was not observed for PM\textsubscript{2.5} constituents, on the contrary, we observed the reverse pattern in correlation showing stronger correlations for short-term exposure (weekly exposures). Correlation between pollutants estimated for the same exposure assessment method was higher for DM, this pattern was also observed in  a previous study \cite{klompmaker2021}. One potential explanation could be that DM models often rely on the same emission source inventory to predict different air pollutants that could lead to higher correlations between different pollutants in comparison to other approaches such as LUR and HM models. PM\textsubscript{2.5} constituents were positive correlated between them for both long- and-short term windows of exposure for LUR and HM models. PM\textsubscript{2.5 - Zn} estimated using LUR and HM models were poorly correlated against estimated pollutants (NO\textsubscript{2}, PM\textsubscript{2.5}, and BC) from other exposure assessment methods, this pattern was observed for long-term and short-term exposure but was more pronounced for HM estimated pollutants. We believe that this results could be partially explained since both PM\textsubscript{2.5 - Fe} and PM\textsubscript{2.5 - Cu} come mainly from traffic, particularly from venhicle non-exhaust factors, while PM\textsubscript{2.5 - Zn} could came from industry emission sources as well this could explain the magnitude of some correlations in \textbf{Fig 5.A} and \textbf{Fig 5.B} \cite{amato2024}. Overall, there was good agreement among the LUR, DM, and HM model estimates for all pollutants. However, the lowest agreement was observed for higher air pollution estimates, suggesting that at elevated pollution levels, discrepancies between model estimates become more pronounced. Our reported findings were consistent with similar studies, which report similar performance metrics for the outdoor air pollutants assessed and relatively good agreement across different \cite{mila2023} and models exposure assessment methods \\

% --- Strengths --- %
Our study has several strengths. First, we use three different approaches to develop outdoor air pollution models for different TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn), allowing for a comprehensive assessment of the effects of air pollution in further epidemiological studies. Secondly, we take into account the potential measurement error introduced by the COVID pandemic in the estimations of our models by incorporating detailed information from pre-and- post pandemic traffic across the city, XVPCA monitoring station network, and traffic scenarios, allowing us to capture changes in air pollution trends across Barcelona. Thirdly, we report different metrics to assess the performance of outdoor air pollution models that are commonly employed to assess performance in different approaches to estimate air pollution levels. Additionally, we report the results for DM models which are frequently overlooked in the exposure assessment components of epidemiological studies. Fourthly, our LUR and HM models had the advantage of a large training dataset in measured concentrations by combining measurements from the BiSC-home and BiSCAPE monitoring campaign with different contrasts in exposure levels across Barcelona, enhancing the robustness and depth of our analysis. Finally, to our knowledge, this is one of the first studies to comprehensively assess different TRAP's using a multifaceted modeling strategy that spans both pre- and post-pandemic periods. It assesses their relative performance at an urban scale, enabling the disentanglement of differences in effect estimates derived from distinct exposure assessment methodologies. This approach allows for a detailed examination of within-city variability in exposure, contributing valuable insights into further epidemiological studies.\\

% --- Limitations --- % 
The results obtained in this study, which stands as one of the first comprehensive assessments of TRAPs, must be interpreted within the context of its limitations. 
First, the methodology employed in the construction of our different models difficult direct comparisons between models, thus making it challenging to draw strong conclusions about their relative performance and accuracy. This inherent complexity in model comparison underscores the need for cautious interpretation of our results. Second, the DM was designed to predict air pollution levels at specific locations and within particular temporal frames. This specificity limits the reproducibility and broader applicability of the HM, which rely on dispersion estimates as a fundamental component of their predictive variables. However, this does not limit the potential utility of our LUR models that could be extended and used in  following epidemiological studies, out of the initial period for what the model was built. These models, despite the study's limitations, offer valuable insights and could be adapted for use in future epidemiological research beyond the initial scope for which they were developed. Third, the absence of external validation data restricts our capacity to evaluate  the accuracy of our preferred modelling approach. Despite this, we assess the performance of the NO\textsubscript{2} model given some lights regarding the most suitable approach for our context. However, we can no extending these findings to other pollutants due to differences in their sources, behavior, and environmental interactions. Despite these challenges, our study lays the groundwork for future research in exposure assessment in urban areas, providing a basis for methodological refinement and encouraging the development of more generalized models that can be  externally validated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Conclusions --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

In conclusion, our study innovatively developed and compared three different exposure assessment methods to estimate outdoor air pollution exposures to several TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn) within the urban context of Barcelona. Our work provides a comprehensive evaluation of outdoor air pollution exposure, highlighting the critical role of diverse exposure assessment methods in understanding the effects and discrepancies of TRAPs exposure. Finally, our exposure estimates derived from the models developed will be used in the BiSC cohort and potentially in further epidemiological studies in Barcelona and its metropolitan area.

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- References --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

\bibliography{references}

\end{document}